---
title: "DESEQ2 Analysis, Tau/Amyloid Mouse Lines"
output: html_notebook
---

The purpose of this notebook is to conduct a differential gene expression analysis with the data obtained from:
Castanho I, Murray TK, Hannon E, Jeffries A et al. Transcriptional Signatures of Tau and Amyloid Neuropathology. Cell Rep 2020 Feb 11;30(6):2040-2054.e5. PMID: 32049030

```{r}

library(dplyr)

library(pheatmap)

library(ggplot2)
```



### Step 1: Downloa files from HPC,read in files into R

Transfer files from HPC to Google drive. The files that are being transferred here are the gene count fiels that were produced by STAR. 

Import the counts for the J20 and rTg4510 mouse line:


```{r}

```

```{r}
#Open the file containing the SRA IDs of the J20 samples and rTg4510 samples
J20_samples<-read.delim('/Users/danielaquijano/Documents/GitHub/Transcriptomics-Final-Project-/Data_Retrieval/SRR_Acc_List_J20_6m_12m.txt',header=FALSE)

```

```{r}
J20_samples[1]
```


```{r}
rTg4510_samples<-
```

```{r}
rTg4510_samples[1]
```

```{r}

sample_names <-cbind(J20_samples[1],)
sample_condition_mouse_model <- c(rep('J20',42),rep('rTg4510',30))
sample_condition_mouse_age<-c(rep('6',4),rep('12',4),rep('12',4))
```

```{r}
#File path Daniela
files_D <- file.path('/Users/danielaquijano/Documents/Final_Project_Files',paste(sample_names,"ReadsPerGene.out.tab",sep=""))
```

```{r}
#File path Tasnim
files_T<-

```

```{r}
files<-rbind(files_D,files_T)
```

```{r}
names(files) <- sample_names
```


Create a matrix of gene counts from the count data outputted by STAR. We need to check whether the library preparation was stranded or unstranded. In order to do this, we can count the number of reads mapped to each strand in a sample with awk:

```{bash}
[dq2033@log-2 SRR8512365]$ grep -v "N_" SRR8512365ReadsPerGene.out.tab | awk '{unst+=$2;forw+=$3;rev+=$4}END{print unst,forw,rev}'
8779853 8870916 314680
```
8870916 Reads 1 (forward) were mapped to known genes and  314680 Reads 2 (reverse) were mapped to known genes. For our analysis, we will use the data coming from the forward strand. 


### Step 2: Join Read Counts from each sample into one matrix

Every sample has its own gene count data in the form of a ReadsPerGene.out.tab file. A sample of what the ReadsPerGene.out.tab file looks like is seen below:

```{r}
read.table('/Users/danielaquijano/Documents/Final_Project_Files/SRR8512351ReadsPerGene.out.tab',sep="\t")
```
As established with the results of the awk command above, the output has to be chosen based on the strandedness of the data. This can be accomplished using the data.table library in order to select the column containing the gene names and mapping information (V1) and also

 select = c("x1", "x2", "x4")

```{r}
install.packages("data.table") # Install data.table
library("data.table")    
```

Loop through the directory of files, open them, add the sample data to the count data matrix
Only Import Selected Variables Using fread() Function of data.table Package. Each sample 
```{r}
i=1
for (file in files) {
  count_data[sample_names[i]]<-read.table(files_D,sep="\t", select = c("x1", "x2", "x4"))
  i=i+1
  
}
```

```{r}
print(files)
```

***** check design for deseq with two conditions *****
```{r}

samples <- data.frame(sample_names=sample_names,condition_a=sample_condition_mouse_model, condition_b=sample_condition_mouse_age)
row.names(samples) <- sample_names
```

