---
title: "R Notebook"
output: html_notebook
---

The goals for this analysis are to investigate how the gene expression changes based on the mouse model, genotype within the mouse model, and age.
Questions to answer:
-What genes are deferentially expressed due to Mouse Model Alone?
-How does aging affect gene expression within each mouse model?
-How does the genotype within each mouse model affect gene expression at different time points?

### Import experimental design matrix.

```{r}
library(DESeq2)
library(pheatmap)
library(dplyr)
library(dendextend)
```

```{r}
design_matrix<-read.table('/Users/tasnimtabassum/Documents/Transcriptomics_SP22/Experimental_Design_TG.csv',sep=',',header=TRUE)

head(design_matrix)
```

```{r}
rownames(design_matrix)<-design_matrix$Sample
design_matrix$Sample<-NULL
```
```{r}
design_matrix
```

```{r}
counts_matrix<-read.table("/Users/tasnimtabassum/Documents/Transcriptomics_SP22/GitHub/Transcriptomics-Final-Project-/Count_Tables/allcounts.csv",sep=',',header=TRUE)
```
```{r}
counts_matrix
```


Because the numbers after the dot in the ensembl IDs represent versions of genes in certain annotations, we can remove these to more easily conduct our differential gene expression analysis. 
```{r}
counts_matrix$V1<-gsub("\\..*","",counts_matrix$V1)
```
```{r}
counts_matrix
```

```{r}
# remove the "V1" from col 1
rownames(counts_matrix)<-counts_matrix$V1
counts_matrix$V1<-NULL

head(counts_matrix)
```
```{r}
counts_matrix<-counts_matrix[,order(colnames(counts_matrix))]
counts_matrix
```

```{r}
design_matrix<-design_matrix[order(rownames(design_matrix)),]

design_matrix

```

```{r}
dds <- DESeqDataSetFromMatrix(countData = counts_matrix,
                              colData = design_matrix,
                              design = ~ Genotype + Age+ Model)
dds

keep <- rowSums(counts(dds)) >= 10
dds <- dds[keep,]
```


```{r}
dds <- DESeq(dds)
```

```{r}
rawcounts.matrix <- counts(dds,normalized=F)
normalizedcounts.matrix <- counts(dds,normalized=T)

```


```{r}
res_1 <- results(dds, contrast = c("Genotype","rtg4510","WT_TG"))

res1_ordered <- res_1[order(res_1$pvalue),] 
head(res1_ordered,200)
```

#Filter Differentially Expressed Genes by p value 0.05 and lfc <-1 and >+1
```{r}
diffexpgenes= data.frame(subset(res1_ordered, padj < 0.05 & (log2FoldChange > 1 | log2FoldChange < -1)))

nrow(diffexpgenes)
```

Install Mouse annotation library:


```{r}
library(biomaRt) #For conversion of transcript IDs to gene ID

library(annotables) #to retrieve grcm38 annotation for mouse genome
library(org.Mm.eg.db) #Mouse genome annotation
library(DOSE)
library(pathview)
library(clusterProfiler)
library(AnnotationHub) 
library(ensembldb)
library(tidyverse)
library(ggnewscale)
```

```{r}
grcm38
idx <- grcm38$ensgene %in% rownames(res1_ordered)
head(idx)
```


```{r}
ids <- grcm38[idx, ]

head(ids)

non_duplicates <- which(duplicated(ids$ensgene) == FALSE)
ids <- ids[non_duplicates, ]

nrow(res1_ordered)
rownames(res1_ordered)
entrezID= grcm38[grcm38$ensgene %in% rownames(res1_ordered), ]

head(entrezID)

entrez_ID_vector = c(entrezID[[2]])

head(entrez_ID_vector)

res2= data.frame(log2foldchange= subset(res1_ordered$log2FoldChange, grcm38$ensgene %in% rownames(res1_ordered)))
padj = subset(res1_ordered$padj, grcm38$ensgene %in% rownames(res1_ordered))
res2 = cbind(padj, res2)
res2 = cbind(entrez_ID_vector, res2)

res2 = na.omit(res2)

head(res2)
#res2 = cbind(res1_ordered$log2FoldChange, res2)

#res1_ordered_tb <- res1_ordered %>%
  #data.frame() %>%
  #rownames_to_column(var="ensgene") %>% 
  #as_tibble()

#res_entrez= as.data.frame(inner_join(res1_ordered_tb, entrezID))


#res_entrez= inner_join(res1_ordered_tb, entrezID, by= c("gene" = ))
```
```{r}
## Create background dataset for hypergeometric testing using all genes tested for significance in the results                 
allOE_genes <- as.character(res_entrez$ensgene)

## Extract significant results
sigOE <- filter(res_entrez, padj < 0.05)

sigOE_genes <- as.character(sigOE$ensgene)
```

```{r}
## Run GO enrichment analysis 
ego <- enrichGO(gene = sigOE_genes, 
                universe = allOE_genes,
                keyType = "ENSEMBL",
                OrgDb = org.Hs.eg.db, 
                ont = "BP", 
                pAdjustMethod = "BH", 
                qvalueCutoff = 0.05, 
                readable = TRUE)
                
## Output results from GO analysis to a table
cluster_summary <- data.frame(ego)
```

```{r}

```

```{r}
```


```{r}
```


```{r}
# Set-up

#BiocManager::install("SPIA")
library(SPIA)

## Significant genes is a vector of fold changes where the names are ENTREZ gene IDs. The background set is a vector of all the genes represented on the platform.

background_entrez <- res2$entrez_ID_vector

sig_res_entrez <- res2[which(res2$padj < 0.05), ]

sig_entrez <- sig_res_entrez$log2foldchange

head(sig_res_entrez)



names(sig_entrez) <- sig_res_entrez$entrez

head(sig_entrez)

dups<-unique(names(sig_entrez[which(duplicated(names(sig_entrez)))]))
sig_entrez<-sig_entrez[!(names(sig_entrez) %in% dups)]
```


```{r}

# currently stuck at this error message
#de= as.vector(sig_entrez)

#de = sort(de, decreasing = FALSE)


spia_result <- spia(de=sig_entrez, all=background_entrez, organism="mmu")

head(spia_result)

```
```{r}
plotP(spia_result, threshold=0.05) 
```
```{r}

write.csv(spia_result, file = "spia_result")
```


```{r}
subset(spia_result, ID == "04727")


```


